Incorporating periodic variability in hidden Markov models for animal movement

### 2016-07-07 paper3.Rnw
failing to compile, I don't know why

\doublespacing
<<pkgs,message=FALSE,warning=FALSE,echo=FALSE>>=
library("ggplot2"); theme_set(theme_bw())
scale_colour_discrete <- function(...,palette="Set1")
    scale_colour_brewer(...,palette=palette)
scale_fill_discrete <- function(...,palette="Set1")
    scale_fill_brewer(...,palette=palette)
zmargin <- theme(panel.margin=grid::unit(0,"lines"))
library("knitr")
opts_chunk$set(fig.width=6,fig.height=4)
library("depmixS4")
library("plyr")
library("dplyr")
library("reshape2")
library("gridExtra")
library("RColorBrewer")
if (!require("ggalt")) {
    cat("installing ggalt from BB repo ...\n")
    library("devtools")
    install_github("bbolker/ggalt")
    library("ggalt")
}
# library("ggalt")
@

@@

@@

Model complexity and the number of parameters increase as the number of latent states increase, FMM to HMM, and lastly, FMM and HMM incorporating temporal heterogeneity (Supplementary material, Table 1).

@@

<<state_plots, warning=FALSE, message=FALSE, echo=FALSE, fig.cap="Step-length parameters (\\textbf{todo: change scales from log() to log-scale; need a real/more complete figure caption here; emphasize that dashes are for emphasis only; rename models (time-homog/heterog)}">>=
g0 <- ggplot(all_cat_pars,aes(log_mean,log_sd,color=Model))+
    geom_point()+geom_line() + facet_wrap(~catID,labeller=label_both)+zmargin
(g0c <- g0 +
    geom_encircle(data=subset(all_cat_pars,
                              (catID=="15" & log_mean<1.5) |
                              (catID!="15" & log_mean<1.8)),
                  colour="black",linetype=2,expand=0.05))
@

@@

<<viterbi, warning=FALSE, message=FALSE, echo=FALSE >>=
load("cat1.RData")
load("cat1sumdat.RData")
vitcat <- function(catt,mod){
  params <- tail(getpars(mod),6)
  vitdist <- numeric(ntimes(mod))
  states <- viterbi(mod)$state
  for(i in 1:ntimes(mod)){
    if(states[i] == 1){
      vitdist[i] <- rnorm(1,params[1],params[2])
    }
    if(states[i] == 2){
      vitdist[i] <- rnorm(1,params[3],params[4])
    }
    if(states[i] == 3){
      vitdist[i] <- rnorm(1,params[5],params[6])
    }
  }
  dat <- data.frame(viterbi=vitdist,time=catt$Time, obs=catt$LogDist)
  return(dat)
}

temp <- vitcat(cat,fithomo3s) %>% dplyr::select(-time)%>%ldply(.,acffun)
ggplot(temp,aes(lag,ACF,colour=.id))+
geom_point()+geom_line() +theme_bw() +
ggtitle("ACF plot of BIC Selected Models")
@

>>>>

\doublespacing
<<pkgs,message=FALSE,warning=FALSE,echo=FALSE>>=
library("ggplot2"); theme_set(theme_bw())
scale_colour_discrete <- function(...,palette="Set1")
    scale_colour_brewer(...,palette=palette)
scale_fill_discrete <- function(...,palette="Set1")
    scale_fill_brewer(...,palette=palette)
zmargin <- theme(panel.margin=grid::unit(0,"lines"))
library("knitr")
opts_chunk$set(fig.width=6,fig.height=4)
library("depmixS4")
library("plyr")
library("dplyr")
library("reshape2")
library("gridExtra")
library("RColorBrewer")
# if (!require("ggalt")) {
#     cat("installing ggalt from BB repo ...\n")
#     library("devtools")
#     install_github("bbolker/ggalt")
#     library("ggalt")
# }
# library("ggalt")
@

@@

Model complexity and the number of parameters increase as the number of latent states increase, FMM to HMM, and lastly, FMM and HMM incorporating temporal heterogeneity (Supplementary material, Table 1). The number of free parameters in an HMM can be generalize by summing up the number of free parameters of the three distinct components:

@@

@@

<<state_plots, warning=FALSE, message=FALSE, echo=FALSE, fig.cap="Step-length parameters (\\textbf{todo: change scales from log() to log-scale; need a real/more complete figure caption here; emphasize that dashes are for emphasis only; rename models (time-homog/heterog)}">>=
# g0 <- ggplot(all_cat_pars,aes(log_mean,log_sd,color=Model))+
#     geom_point()+geom_line() + facet_wrap(~catID,labeller=label_both)+zmargin
# (g0c <- g0 +
#     geom_encircle(data=subset(all_cat_pars,
#                               (catID=="15" & log_mean<1.5) |
#                               (catID!="15" & log_mean<1.8)),
#                   colour="black",linetype=2,expand=0.05))
@

@@
